---
title: "notebook_project"
format: html
editor: visual
---

```{r}
# load libraries 

library(tidyverse)
library(janitor)
library(ggplot2)
library(dplyr)
library(ggthemes)

```

## Explanation of notebook:

# The goal of this notebook is to analyze the correlation between average incomes and internet access in Baltimore City neighborhoods. To do this I am using a dataset that includes the median income in Baltimore City neighborhoods based on 2020 Census data. I am also using an open source dataset from the City of Baltimore that details how many homes do not have internet subscriptions in each Baltimore City neighborhood.

## Questions I am asking the data

# Do high income areas have high internet access?

# Do low income areas have poor internet access?

# Does internet access improve/get worse for places whose median income increased/decreased 2016-2020?

## Findings

# While neighborhoods with good internet access usually have higher median incomes and neighborhoods with low median incomes usually have poor internet access, there does not seem to be a significant association between decreasing median incomes and lower internet access overtime. Most neighborhoods increase internet access even when they saw negative percent change in median income.

## Caveats in the data

# in both datasets, some neighborhoods are repeated multiple times with different numbers in the same year, so it is unclear which number is the most accurate representation of the neighborhood. The internet subscription dataset also does not detail how they define an internet subscription(ie. broadband or wireless).

## Author

# Victoria Stavish for Jour472

```{r}
# load income data

balt_income <- read_csv("baltcity_income_clean.csv")%>%
clean_names()

## Data dictionary
# This dataset has 6 columns. The first column is the names of Baltimore City neighborhoods (chr), the following three columns show the median income in each neighborhood in 2010 (dbl), 2016 (dbl) and 2020 (dbl). The fourth column shows the census id for each neighborhood (dbl) and the fifth column shows the geography id of each neighborhood (dbl). 

# load internet data

balt_internet <- read_csv("balt_internet.csv") %>%
  clean_names()

## Data dictionary
# This dataset has 9 columns. The first column gives an object id (dbl) to each row. The second column gives the Baltimore City neighborhood name as of the 2010 census data (chr). The following three columns give the percentage of households in each neighborhood that do not have an internet subscription for the years 2017 (dbl), 2018 (dbl) and 2019 (dbl). The sixth column gives the Baltimore City neighborhood name as of the 2020 census data (chr). The seventh column gives the percentage of households in each neighborhood that do not have an internet subscription for the year 2020 (dbl). The eighth and ninth columns give the neighborhood shape (dbl) and length (dbl), neither of which are relevant for our analyses. 

```

# Do low income neighborhoods have low percentage of internet access?

# The ten neighborhoods with highest percentage of homes with no internet access, what is the median income?

```{r}

# 10 neighborhoods with lowest internet access

balt_internet_b10 <- balt_internet %>%
  as.data.frame() %>%
  clean_names() %>%
 dplyr::: select(csa2020, nohhint20) %>%
  arrange(desc(nohhint20)) %>%
  slice_max(nohhint20, n=10)

# median percentage of homes without internet access among the 10 neighborhoods with the lowest internet access = 31.66%
summary(balt_internet_b10)

# in neighborhoods with lowest percentage of internet access, what is the median salary?

b10_internet_income <- filter(balt_income, neighborhood %in% c("Sandtown-Winchester/Harlem Park", "Greater Rosemont", "Forest Park/Walbrook", "Greater Mondawmin", "Pimlico/Arlington/Hilltop", "Cherry Hill", "Brooklyn/Curtis Bay/Hawkins Pointl", "Oliver/Johnson Square", "Southwest Baltimore", "Southern Park Heights")) %>%
  dplyr:::select(neighborhood, x2020) %>%
  group_by(neighborhood) %>% 
  summarize(mean(x2020))

# median income in neighborhoods with lowest percentage of internet access = $33,967
summary(b10_internet_income)
  
```

# Do high income neighborhoods have high percentage of internet access?

# The ten neighborhoods with highest percentage of homes with internet access, what is the median income?

```{r}

# 10 neighborhoods with high internet access

t10_balt_internet <- balt_internet %>%
  as.data.frame() %>%
  clean_names() %>%
 dplyr::: select(csa2020, nohhint20) %>%
  arrange(desc(nohhint20)) %>%
  slice_min(nohhint20, n=10)

# median percentage of homes without internet access among the 10 neighborhoods with high internet access = 5.668% 
summary(t10_balt_internet)

# in neighborhoods with highest internet access, what is the median income?

t10_internet_income <- filter(balt_income, neighborhood %in% c("Greater Roland Park/Poplar Hill", "South Baltimore", "Canton", "	
Downtown/Seton Hill", "Fells Point", "Greater Lauraville", "North Baltimore/Guilford/Homeland", "	
Chinquapin Park/Belvedere", "	
Inner Harbor/Federal Hill", "Hamilton")) %>%
  dplyr:::select(neighborhood, x2020) %>%
  group_by(neighborhood) %>% 
  summarize(mean(x2020))

# median income in neighborhoods with highest percentage of internet access = 100,005
summary(t10_internet_income)

```

# Does internet access improve/get worse for places whose median income increased/decreased 2016-2020?

```{r}
# top 10 neighborhoods with greatest growth in median income 2016-2020

poschange_income <- balt_income %>%
  clean_names() %>%
  dplyr:::select(neighborhood, x2016, x2020) %>%
  mutate(pct_change = ((x2020 - x2016)/x2016*100)) %>%
  arrange(desc(pct_change)) %>%
  slice_max(pct_change, n=10)

# Those 10 neighborhoods, how did their internet access change between 2017-2020

poschange_internet <- filter(balt_internet, csa2020 %in% c("Greater Charles Village/Barclay", "Clifton-Berea", "Allendale/Irvington/S. Hilton", "Poppleton/The Terraces/Hollins Market", "Orangeville/East Highlandtown", "The Waverlies", "	
Morrell Park/Violetville", "Upton/Druid Heights", "	
Southwest Baltimore", "Greenmount East")) %>%
  dplyr:::select(csa2020, nohhint17, nohhint20) %>%
  mutate(change = (nohhint20 - nohhint17))

```
 
# A general look at changes in internet access compared to changes in median income in Baltimore City neighborhoods 2016-2020.
 
```{r}
income_change <- balt_income %>%
  dplyr:::select(neighborhood, x2016, x2020) %>%
  mutate(income_pct_change = ((x2020 - x2016)/x2016*100)) 


# mutate the internet data to see percent change over time

internet_change <- balt_internet %>%
  clean_names() %>%
  dplyr:::select(csa2020, nohhint17, nohhint20) %>%
  mutate(internet_pct_change = ((nohhint20 - nohhint17)/nohhint17*100)) 
  
# join the change in internet dataframe with the change in income dataframe
  
change_join <- internet_change %>%
  inner_join(income_change, c("csa2020" = "neighborhood"))


```

# Does internet access improve/get worse for places whose median income increased/decreased 2016-2020?

```{r}

# 10 neighborhoods with greatest decrease in median income 2016-2020

negchange_income <- balt_income %>%
  clean_names() %>%
  dplyr:::select(neighborhood, x2016, x2020) %>%
  mutate(pct_change = ((x2020 - x2016)/x2016*100)) %>%
  arrange(desc(pct_change)) %>%
  slice_min(pct_change, n=10)

# Those 10 neighborhoods, how did their internet access change between 2017-2020

negchange_internet <- filter(balt_internet, csa2020 %in% c("Southwest Baltimore", "Harbor East/Little Italy", "Greenmount East", "Brooklyn/Curtis Bay/Hawkins Point", "Sandtown-Winchester/Harlem Park", "	
Westport/Mount Winans/Lakeland", "Cross-Country/Cheswolde", "Upton/Druid Heights")) %>%
  dplyr:::select(csa2020, nohhint17, nohhint20) %>%
  mutate(change = (nohhint20 - nohhint17))

```


# Scatterplot of internet access and income in neighborhoods in 2020
```{r}
# Join the dataframes for Baltimore City income and Baltimore City internet

income_internet_join <- balt_income %>%
  clean_names() %>%
  group_by(neighborhood, x2020) %>%
  inner_join(balt_internet, c("neighborhood" = "csa2020")) %>%
  select(neighborhood, x2020, nohhint20)

# Create a scatterplot to show the relationship between income and internet access in Baltimore City neighborhoods

ggplot(income_internet_join, aes(y=nohhint20, x=x2020)) +
geom_point() +
theme_excel() +
  labs(
    title="Internet access and income in Baltimore City neighborhoods",
    subtitle = "In 2020 Baltimore City neighborhoods with higher median incomes generally
had more home with internet subscriptions than neighborhoods with 
low median incomes. Each point represents a neighborhood in Baltimore City,",
    x = "Median income",
    y = "Percent of homes without internet subscriptions",
    caption = "Source: The U.S. Census Bureau  By: Victoria Stavish")

```
# story memo

```{r}
## Armed with your data analysis, write a 300 word reporting memo and story pitch citing your data findings and any relevant background research to provide context. Interviews are not necessary for this assignment but you're welcome to conduct them.

# I am pitching a story on how even though some Baltimore City neighborhood's median income are decreasing, internet access continues to improve among all Baltimore City neighborhoods. For example, the ten Baltimore neighborhoods with the no gains or the smallest gains in median income between 2016 and 2020, all but one saw significant decreases in the percentage of homes without internet access. For example, even though Southwest Baltimore saw a 38% decrease in median income, their percent of households without internet subscriptions decreased by 9% between 2017 and 2020. Upton/Druid Heights also saw their median income decrease by 17% between 2016 and 2020, but their percent of homes without internet also decreased by nearly 17% between 2017 and 2020. While income clearly has an affect, given the scatterplot in this notebook, why does a decrease in median income not necessarily mean a decrease in internet access? Are we seeing a shift towards low income correlating less and less with internet access?

# It is notable that there are some places where both median income increased and the number of households that do not have internet access increased, however, the changes in those numbers are not as drastic and there are not many examples of that, making it less noteworthy than neighborhoods in which internet access increased while median income decreased. 

# However, the general theme of internet access and income having a strong correlation is still true. In 2020, of the 10 neighborhoods in Baltimore City with the best internet access, the median income was about $100,000, while in the neighborhoods in Baltimore City with the worst internet access, the median income was about $34,000. 



## Draft two Tweets and include images (R graphics) to accompany the story.

# tweet 1: While there is a high correlation between internet access and income (as seen in the chart below) we found that some neighborhoods in Baltimore City are seeing gains in internet access even as median income decreases. Read more here: *story link*

# tweet 2: Internet access is improving in some Baltimore City neighborhoods, despite a decreasing median income. We looked at why this is and what it means for the future of internet access in Baltimore. *story link*



```

