---
title: "notebook_project"
format: html
editor: visual
---

```{r}
# load libraries 

library(tidyverse)
library(janitor)
library(ggplot2)
library(dplyr)

```

## Explanation of notebook:

# The goal of this notebook is to analyze the correlation between average incomes and internet access in Baltimore City neighborhoods. To do this I am using a dataset that includes the median income in Baltimore City neighborhoods based on 2020 Census data. I am also using an open source dataset from the City of Baltimore that details how many homes do not have internet subscriptions in each Baltimore City neighborhood. 

## Questions I am asking the data

# Do high income areas have high internet access?

# Do low income areas have poor internet access?

# Does internet access improve/get worse for places whose median income increased/decreased 2016-2020?

## Findings

# While neighborhoods with good internet access usually have higher median incomes and neighborhoods with low median incomes usually have poor internet access, there does not seem to be a significant association between decreasing median incomes and lower internet access overtime. Most neighborhoods increase internet access even when they saw negative percent change in median income.

## Caveats in the data

# in both datasets, some neighborhoods are repeated multiple times with different numbers in the same year, so it is unclear which number is the most accurate representation of the neighborhood. The internet subscription dataset also does not detail how they define an internet subscription(ie. broadband or wireless).

## Author

# Victoria Stavish for Jour472

```{r}
# load income data

balt_income <- read_csv("baltcity_income_clean.csv")%>%
clean_names()

## Data dictionary
# This dataset has 6 columns. The first column is the names of Baltimore City neighborhoods (chr), the following three columns show the median income in each neighborhood in 2010 (dbl), 2016 (dbl) and 2020 (dbl). The fourth column shows the census id for each neighborhood (dbl) and the fifth column shows the geography id of each neighborhood (dbl). 

# load internet data

balt_internet <- read_csv("balt_internet.csv") %>%
  clean_names()

## Data dictionary
# This dataset has 9 columns. The first column gives an object id (dbl) to each row. The second column gives the Baltimore City neighborhood name as of the 2010 census data (chr). The following three columns give the percentage of households in each neighborhood that do not have an internet subscription for the years 2017 (dbl), 2018 (dbl) and 2019 (dbl). The sixth column gives the Baltimore City neighborhood name as of the 2020 census data (chr). The seventh column gives the percentage of households in each neighborhood that do not have an internet subscription for the year 2020 (dbl). The eighth and ninth columns give the neighborhood shape (dbl) and length (dbl), neither of which are relevant for our analyses. 

```

# Do low income neighborhoods have low percentage of internet access?

# The ten neighborhoods with highest percentage of homes with no internet access, what is the median income?

```{r}

# 10 neighborhoods with lowest internet access

balt_internet_b10 <- balt_internet %>%
  as.data.frame() %>%
  clean_names() %>%
 dplyr::: select(csa2020, nohhint20) %>%
  arrange(desc(nohhint20)) %>%
  slice_max(nohhint20, n=10)

# median percentage of homes without internet access among the 10 neighborhoods with the lowest internet access = 31.66%
summary(balt_internet_b10)

# in neighborhoods with lowest percentage of internet access, what is the median salary?

b10_internet_income <- filter(balt_income, neighborhood %in% c("Sandtown-Winchester/Harlem Park", "Greater Rosemont", "Forest Park/Walbrook", "Greater Mondawmin", "Pimlico/Arlington/Hilltop", "Cherry Hill", "Brooklyn/Curtis Bay/Hawkins Pointl", "Oliver/Johnson Square", "Southwest Baltimore", "Southern Park Heights")) %>%
  dplyr:::select(neighborhood, x2020) %>%
  group_by(neighborhood) %>% 
  summarize(mean(x2020)

# median income in neighborhoods with lowest percentage of internet access = $33,967
summary(b10_internet_income)
  
```

# Do high income neighborhoods have high percentage of internet access?

# The ten neighborhoods with highest percentage of homes with internet access, what is the median income?

```{r}

# 10 neighborhoods with high internet access

t10_balt_internet <- balt_internet %>%
  as.data.frame() %>%
  clean_names() %>%
 dplyr::: select(csa2020, nohhint20) %>%
  arrange(desc(nohhint20)) %>%
  slice_min(nohhint20, n=10)

# median percentage of homes without internet access among the 10 neighborhoods with high internet access = 5.668% 
summary(t10_balt_internet)

# in neighborhoods with highest internet access, what is the median income?

t10_internet_income <- filter(balt_income, neighborhood %in% c("Greater Roland Park/Poplar Hill", "South Baltimore", "Canton", "	
Downtown/Seton Hill", "Fells Point", "Greater Lauraville", "North Baltimore/Guilford/Homeland", "	
Chinquapin Park/Belvedere", "	
Inner Harbor/Federal Hill", "Hamilton")) %>%
  dplyr:::select(neighborhood, x2020)

# median income in neighborhoods with highest percentage of internet access = 100,005
summary(t10_internet_income)

```

# Does internet access improve/get worse for places whose median income increased/decreased 2016-2020?

```{r}
# top 10 neighborhoods with greatest growth in median income 2016-2020

poschange_income <- balt_income %>%
  clean_names() %>%
  dplyr:::select(neighborhood, x2016, x2020) %>%
  mutate(pct_change = ((x2020 - x2016)/x2016*100)) %>%
  arrange(desc(pct_change)) %>%
  slice_max(pct_change, n=10)

# Those 10 neighborhoods, how did their internet access change between 2017-2020

poschange_internet <- filter(balt_internet, csa2020 %in% c("Greater Charles Village/Barclay", "Clifton-Berea", "Allendale/Irvington/S. Hilton", "Poppleton/The Terraces/Hollins Market", "Orangeville/East Highlandtown", "The Waverlies", "	
Morrell Park/Violetville", "Upton/Druid Heights", "	
Southwest Baltimore", "Greenmount East")) %>%
  dplyr:::select(csa2020, nohhint17, nohhint20) %>%
  mutate(change = (nohhint20 - nohhint17))

```

# Does internet access improve/get worse for places whose median income increased/decreased 2016-2020?

```{r}

# 10 neighborhoods with greatest decrease in median income 2016-2020

negchange_income <- balt_income %>%
  clean_names() %>%
  dplyr:::select(neighborhood, x2016, x2020) %>%
  mutate(pct_change = ((x2020 - x2016)/x2016*100)) %>%
  arrange(desc(pct_change)) %>%
  slice_min(pct_change, n=10)

# Those 10 neighborhoods, how did their internet access change between 2017-2020

negchange_internet <- filter(balt_internet, csa2020 %in% c("Southwest Baltimore", "Harbor East/Little Italy", "Greenmount East", "Brooklyn/Curtis Bay/Hawkins Point", "Sandtown-Winchester/Harlem Park", "	
Westport/Mount Winans/Lakeland", "Cross-Country/Cheswolde", "Upton/Druid Heights")) %>%
  dplyr:::select(csa2020, nohhint17, nohhint20) %>%
  mutate(change = (nohhint20 - nohhint17))


```
