---
title: "notebook_project"
format: html
editor: visual
---

```{r}
# load libraries 

library(tidyverse)
library(janitor)
library(ggplot2)
library(dplyr)

```

## Explanation of notebook:

# The goal of this notebook is to analyze the correlation between average incomes and internet access in Baltimore City neighborhoods. To do this I am using a dataset that includes the median income in Baltimore City neighborhoods based on 2020 Census data. I am also using an open source dataset from the City of Baltimore that details how many homes do not have internet subscriptions in each Baltimore City neighborhood.

## Questions I am asking the data

# Do high income areas have high internet access?

# Do low income areas have poor internet access?

# Does internet access improve/get worse for places whose median income increased/decreased 2016-2020?

## Findings

# While neighborhoods with good internet access usually have higher median incomes and neighborhoods with low median incomes usually have poor internet access, there does not seem to be a significant association between decreasing median incomes and lower internet access overtime. Most neighborhoods increase internet access even when they saw negative percent change in median income. 

## Caveats in the data

# in both datasets, some neighborhoods are repeated multiple times with different numbers in the same year, so it is unclear which number is the most accurate representation of the neighborhood.

## Author

# Victoria Stavish for Jour472

```{r}
# load income data

balt_income <- read_csv("baltcity_income_clean.csv")%>%
clean_names()

# load internet data

balt_internet <- read_csv("balt_internet.csv") %>%
  clean_names()

```

# Do low income neighborhoods have low percentage of internet access?

# The ten neighborhoods with highest percentage of homes with no internet access, what is the median income?

```{r}

# 10 neighborhoods with lowest internet access

balt_internet_b10 <- balt_internet %>%
  as.data.frame() %>%
  clean_names() %>%
 dplyr::: select(csa2020, nohhint20) %>%
  arrange(desc(nohhint20)) %>%
  slice_max(nohhint20, n=10)

# median percentage of homes without internet access among the 10 neighborhoods with the lowest internet access = 31.66%
summary(balt_internet_b10)

# in neighborhoods with lowest percentage of internet access, what is the median salary?

b10_internet_income <- filter(balt_income, neighborhood %in% c("Sandtown-Winchester/Harlem Park", "Greater Rosemont", "Forest Park/Walbrook", "Greater Mondawmin", "Pimlico/Arlington/Hilltop", "Cherry Hill", "Brooklyn/Curtis Bay/Hawkins Pointl", "Oliver/Johnson Square", "Southwest Baltimore", "Southern Park Heights")) %>%
  dplyr:::select(neighborhood, x2020)

# median income in neighborhoods with lowest percentage of internet access = 33,967
summary(b10_internet_income)
  
```
# Do high income neighborhoods have high percentage of internet access?

# The ten neighborhoods with highest percentage of homes with internet access, what is the median income?
```{r}

# 10 neighborhoods with high internet access

t10_balt_internet <- balt_internet %>%
  as.data.frame() %>%
  clean_names() %>%
 dplyr::: select(csa2020, nohhint20) %>%
  arrange(desc(nohhint20)) %>%
  slice_min(nohhint20, n=10)

# median percentage of homes without internet access among the 10 neighborhoods with high internet access = 5.668% 
summary(t10_balt_internet)

# in neighborhoods with highest internet access, what is the median income?

t10_internet_income <- filter(balt_income, neighborhood %in% c("Greater Roland Park/Poplar Hill", "South Baltimore", "Canton", "	
Downtown/Seton Hill", "Fells Point", "Greater Lauraville", "North Baltimore/Guilford/Homeland", "	
Chinquapin Park/Belvedere", "	
Inner Harbor/Federal Hill", "Hamilton")) %>%
  dplyr:::select(neighborhood, x2020)

# median income in neighborhoods with highest percentage of internet access = 100,005
summary(t10_internet_income)

```
# Does internet access improve/get worse for places whose median income increased/decreased 2016-2020?

```{r}
# top 10 neighborhoods with greatest growth in median income 2016-2020

poschange_income <- balt_income %>%
  clean_names() %>%
  dplyr:::select(neighborhood, x2016, x2020) %>%
  mutate(pct_change = ((x2020 - x2016)/x2016*100)) %>%
  arrange(desc(pct_change)) %>%
  slice_max(pct_change, n=10)

# Those 10 neighborhoods, how did their internet access change between 2017-2020

poschange_internet <- filter(balt_internet, csa2020 %in% c("Greater Charles Village/Barclay", "Clifton-Berea", "Allendale/Irvington/S. Hilton", "Poppleton/The Terraces/Hollins Market", "Orangeville/East Highlandtown", "The Waverlies", "	
Morrell Park/Violetville", "Upton/Druid Heights", "	
Southwest Baltimore", "Greenmount East")) %>%
  dplyr:::select(csa2020, nohhint17, nohhint20) %>%
  mutate(change = (nohhint20 - nohhint17))

```

# Does internet access improve/get worse for places whose median income increased/decreased 2016-2020?

```{r}

# 10 neighborhoods with greatest decrease in median income 2016-2020

negchange_income <- balt_income %>%
  clean_names() %>%
  dplyr:::select(neighborhood, x2016, x2020) %>%
  mutate(pct_change = ((x2020 - x2016)/x2016*100)) %>%
  arrange(desc(pct_change)) %>%
  slice_min(pct_change, n=10)

# Those 10 neighborhoods, how did their internet access change between 2017-2020

negchange_internet <- filter(balt_internet, csa2020 %in% c("Southwest Baltimore", "Harbor East/Little Italy", "Greenmount East", "Brooklyn/Curtis Bay/Hawkins Point", "Sandtown-Winchester/Harlem Park", "	
Westport/Mount Winans/Lakeland", "Cross-Country/Cheswolde", "Upton/Druid Heights")) %>%
  dplyr:::select(csa2020, nohhint17, nohhint20) %>%
  mutate(change = (nohhint20 - nohhint17))


```