#Lab_03 {lab_03}

rob wells, 8/14/2021

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
library(tidyverse)

# Load janitor here
library(janitor)

install.packages("dplyr")
library(dplyr)

```

### Load data

In this data set, each row represents a city or town in Maryland with the 2010 and 2020 figures for: 

-  Total population 
-  White only population
-  Black or African American only population
-  Hispanic only population 
-  Native American only population
-  Asian only population
-  Pacific Islander (including native Hawiian) only population

Background on this data from [Social Explorer](https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B02001) 
```{r}
# Write code to load the Maryland race by city, place for 2010 - 2020 and call it md_race.
# The data is called lab_03.csv, and is in the same folder as lab_03.qmd (this data notebook)


# your answer here
md_data <- rio::import("lab_03.csv")

```


## Answer questions

Q1: Which five cities have the largest overall Hispanic population in 2020? Name the five cities that had the largest Hispanic population in 2020. 

A1:
```{r}

# Write your code in this codeblock.

h_pop <- md_data %>%
  select(place, x2020_hispanic) %>%
  arrange(desc(x2020_hispanic))
 

# your answer here
# Baltimore City, Wheaton CDP, Germantown CDP, Silver Spring CDP, Aspen Hill CDP

```

Q2: Find the cities with the largest Black population and arrange the results in descending order. 

A2:

```{r}
#write your code here

b_pop <- md_data %>%
  select(place, x2020_black) %>%
  arrange(desc(x2020_black))

# your answer here
# Baltimore City, Waldorf CDP, Bowie city, Clinton CDP, Columbia CDP

```

Q3: Create a table called top_hispanic that shows the top 15 communities with the largest Hispanic population in 2020, and also display the 2010 Hispanic population and the total community population for 2020. Look at your results - what stands out to you? Craft a one-paragraph story pitch for your editor. 

A3:

```{r}

# your answer here

top_hispanic <- md_data %>%
  select(place, x2010_hispanic, x2020_hispanic, x2020_total) %>%
  arrange(desc(x2020_hispanic))
  head(top_hispanic, 15) 
 
top_hispanic <- md_data %>%
  select(place, x2010_hispanic, x2020_hispanic, x2020_total) %>%
  drop_na(place, x2010_hispanic, x2020_hispanic, x2020_total) %>%
  arrange(desc(x2020_hispanic))
  tail(top_hispanic)
  length(top_hispanic)
  
  top_hispanic %>%
    arrange(top_hispanic, x2020_total, x2020_hispanic, -x2010_hispanic, place)

  
  # Pitch:
  # Why did some Maryland cities see so much more growth in the Hispanic population than others between 2010 and 2020? Baltimore city grew by about 10,000, Aspen Hill CDP grew by about 8,000 and Montgomery Village CDP grew by about 5,000 between 2010 and 2020. Most of the other Maryland cities that currently have the highest Hispanic population only saw growth of around 2,000 and 3,000. What is making certain cities more desirable for Hispanic people than others? And on the contrary, what are the Maryland cities that saw the largest decrease in their Hispanic population? What makes those cities so undesirable?

```
