#lab_07 
Sean Mussenden, Aug 16, 2022

## About this lab

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Make a code block and load janitor and the tidyverse

```{r}
#load libraries

library(tidyverse)
library(janitor)
library(stringr)
```


## Load data

For today's lab, we are using the DEA's ARCOS database, which tracks shipments of opioids and other controlled substances from manufacturers to pharmacies.  As we will read this week, the Washington Post used this data to show which communities were flooded by opioids during the death and overdose crisis of the last two decades. 

There are four tables in the data folder. 

* arcos.csv has the total number of opioid pills shipped to pharmacies located in each U.S. county each year (from 2006-2014). Each row represents one county in one year during that period. In some states, counties and cities are represented. You will need this for both questions.
* county_population_2014.csv has the total population of each U.S. county (or city in some states) in 2014 from the American Community Survey. You will need to use this for the first question.
* state_population_2014.csv has the total population of each U.S. state in 2014 from the American Community Survey. You will need to use this for the second question.
* state_crosswalk.csv is a simple dataframe with two columns listing the state and its abbreviation.  You may need to use this for the second question.

**Task**: Create a codeblock and load the four dataframes using appropriate names below.

```{r}
#load dataframes

shipments <- read_csv("data/arcos.csv")

state_abbr <- read_csv("data/state_crosswalk.csv")

state_pop <- read_csv("data/state_population_2014.csv")

county_pop <- read_csv("data/county_population_2014.csv")

```


## Questions
**Task** Create codeblocks in the space below and answer the questions. 
**Q1** Which county or city received the most total opioid pills per person? Add up all the pills between 2006 and 2014 when doing this calculation. How many total pills, and what was the per person rate?

```{r}
# Q1: Which county or city received the most total opioid pills per person? Add up all the pills between 2006 and 2014 when doing this calculation. How many total pills, and what was the per person rate?

# Group and summarise to find total pill shipments per county. Join shipments and county population dataframes on geoid/countyfips and the  divide total pills by the county population. 

newcountytable <- shipments %>%
  group_by(countyfips, buyer_county, buyer_state) %>%
  summarise(total_pills = sum(total_pills)) %>%
  inner_join(county_pop, c("countyfips" = "geoid")) %>%
  mutate(pills_per_person = (total_pills / total_population_2014)*100) %>%
  arrange(desc(pills_per_person))

# Norton City in Virginia had 262,893.40 pills per person between 2006-2014 and the county had a total shipment number of 10,555,170 between 2006-2014

```


**Q2** Which U.S. state had the highest number of opioid pills per person for pills shipped in 2014? How many total pills, and what was the per person rate? How did that compare to Washington, D.C.?

```{r}
# Q2: Which U.S. state had the highest number of opioid pills per person for pills shipped in 2014? How many total pills, and what was the per person rate? How did that compare to Washington, D.C.?

# make sure the strings representing DC are consistent throughout dataframes

state_pop$state <- str_replace(state_pop$state, "District of Columbia", "Washington, DC")

# create a new dataframe by inner joining state population with state abbreviations so that we can later join and group on state abbreviations

newstate_pop <- state_pop %>%
inner_join(state_abbr, c("state" = "state_name"))

# join newstate_pop with shipments, filter for the year 2014, then group and summarise to get total pills for state and then mutate a column to show pills per person

pills_2014 <- shipments %>%
  select(buyer_county:total_pills) %>%
  inner_join(newstate_pop, c("buyer_state" = "state_abb")) %>%
  filter(year == "2014") %>%
  group_by(buyer_state, total_population_2014) %>%
  summarise(total_pills = sum(total_pills)) %>%
  mutate(pills_per_person = (total_pills / total_population_2014)*100) %>%
  arrange(desc(pills_per_person))
  
# West Virginia had the highest number of opioid pills per person for pills shipped in 2014 at 6,513.188. The total number of pills shipped to West Virginia in 2014 was 120,746,760. In comparison, DC only had 1,356.939 pills per person and 8,599,410 pills total. West Virginia had over four times the amount of pills as DC. 

```


**Q3** Write a sentence that could be dropped in as the lede of a news story from the finding in question two. 

```{r}
# Q3: Write a sentence that could be dropped in as the lede of a news story from the finding in question two. 

# West Virginia was had six thousand opiods delivered to the state per person in 2014, the highest rate by over 200 that year. 

6513.188 - 6252.980


```

