#lab_08
Sean Mussenden, Aug 16, 2022

## About this lab

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load janitor and the tidyverse and sf

```{r}

library(janitor)
library(sf)
library(tidyverse)

```


## Load data

You have been given two spatial dataframes and one regular dataframe to load.  The dataframes are:

* baltimore_homicides.rds | spatial | location and details about each Baltimore homicide between 2007 and 2017, including information about the status (open, closed with arrest, closed without arrest) and the exact location. Details on the data: https://github.com/washingtonpost/data-homicides
* baltimore_tract_shapefiles.rds | spatial | polygons of each census tract in Baltimore
* baltimore_tract_demographics.rds | regular | the total population and white population of each baltimore Census tract in 2017.

**Task** Load the data

```{r}

bmore_homicides <- read_rds("data/baltimore_homicides.rds")
bmore_demographics <- read_rds("data/baltimore_tract_demographics.rds")
bmore_shape <- read_rds("data/baltimore_tract_shapefiles.rds")

```


## Questions

**Q1** Are homicides that occur in majority white Census tracts in Baltimore solved at a higher rate than homicides in non-majority white Baltimore census tracts?
* Write up your findings in three paragraphs that could be dropped into a news story, with the answer to the question as the lede. 
* Include any other relevant details.
* Output a table with relevant results. 
* Output at least one map visualization pertinent to your finding. 

**A1** 

```{r}
majority_w <- bmore_demographics %>%
  clean_names() %>%
  mutate(pct_w = white_pop/total_pop) %>%
  select(geoid,name,pct_w) %>%
  mutate(neighborhood_type = case_when(
    pct_w > .50 ~ "white_majority",
    TRUE ~ "not_white_majority"
  ))
  
majority_w <- majority_w %>%
  as.data.frame() %>%
  inner_join(bmore_shape) %>%
  st_as_sf()

shootings_by_neighborhood <- bmore_homicides %>%
  st_join(majority_w) %>%
  as_tibble()

case_by_neighborhood <- shootings_by_neighborhood %>%
  group_by(neighborhood_type,disposition) %>%
  count() %>%
  pivot_wider(names_from=disposition, values_from=n) %>%
  clean_names() %>%
  mutate(total_shootings = closed_by_arrest + closed_without_arrest + open_no_arrest) %>%
  mutate(unsolved = open_no_arrest + closed_without_arrest) %>%
  mutate(pct_unsolved = unsolved/total_shootings*100)

ggplot() +
  geom_sf(data=bmore_shape) +
  geom_sf(data=bmore_homicides)
```
# Majority White neighborhoods in baltimore have a higher rate of solved homicide cases than majority non-white neighborhoods, according to Census data. 

# Majority white neighborhoods in Baltimore had 46% of their homicides unsolved while majority non-white neighborhoods had about 66% of their homicides unsolved. While non-white majority neighborhoods also had more homicides than majority white neighborhoods in general - non-white majority neighborhoods had 2,588 total shootings while majority white neighborhoods had 239 - majority white neighborhoods still have better crime response overall. 

# Majority white neighborhoods only had 15 homicide closed without arrest while majority non-white neighborhoods had 137 homicide cases closed without arrest. Majority white neighborhoods only had 97 homicide cases open without arrest while majority non-white neighborhoods had 1,577 homicide cases open without arrest. Overall, majority white neighborhoods only have 111 unsolved homicide cases and majority non-whote neighborhoods have 1,714. 


